{\bf {\Large Improving estimates of ecosystem metabolism by reducing effects of tidal advection on dissolved oxygen time series}}

### 2015-01-23 swmp_noise.tex
finished revisions from Hagy comments after Smith

The proportion of daily integrated metabolism estimates that were anomalous (negative production, positive respiration) were significantly reduced for all sites after filtering (\cref{tab:case_res}), perhaps indicating the relative effects of water movement.  Before filtering, anomalous values ranged from 0.09 (as a proportion of the total estimates, Rookery Bay) to 0.22 (Padilla Bay) for production and 0.08 (Rookery Bay) to 0.21 (Elkhorn Slough) for respiration. Anomalous values were reduced to near zero for all case studies, particularly Rookery Bay and Sapelo Island.  Metabolism estimates using filtered \ac{DO} time series had decreased mean production (-63.1 \% change from the annual mean) and respiration (-62.6 \%) for Elkhorn Slough, increased mean production (17.8 \%) and respiration (18.8 \%) for Padilla Bay, and generally unchanged mean production and respiration for Rookery Bay and Sapelo Island (\cref{tab:case_res}).  Changes in mean estimates based on filtered \ac{DO} time series suggests that the weighted regression removed variation attributed to both biological and physical processes.  The implications of these undesirable results are described below.  Decreases in the standard erorr for all metabolism estimates were observed for all cases after filtering.  However, reduction in standard error may also occur if variation from biological processes is reduced individually or in addition to physical variation.  The extent to which this reduction in variation is related to the former should be minimal, provided that the two are statistically distinguishable.  Situations when the phasing of the tidal and solar cycles are correlated could be instances when the two are unable to be separated by the model.  Reductions in mean annual values (particularly at Elkhorn Slough) in addition to standard errors suggests that biological variation may have been removed.

@@

% case study metabolism results, including perc anom
%latex.default(tab, file = "", rowlabel = "Site", insert.bottom = foot.val,     caption = cap.val, caption.loc = "top", rgroup = unique(to_tab$Site),     n.rgroup = rep(2, 4), cgroup = c("Pg\\textsuperscript{\\textit{a}}",         "Rt"), n.cgroup = c(3, 3), rowname = rows, colheads = c(rep(c("Mean",         "Std. Err.", "Anom"), 2), c("Mean", "Std. Err.")), label = "tab:case_res")%
\begin{table}[!tbp]
\caption{Summary of metabolism estimates (gross production, respiration) for case studies using \ac{DO} time series before (observed) and after (filtered) filtering with weighted regression.  Means and standard errors are based on daily integrated metabolism estimates.  Anomalous values are the proportion of metabolism estimates that were negative for gross production and positive for respiration.\label{tab:case_res}} 
\begin{center}
\begin{tabular}{llllclll}
\hline\hline
\multicolumn{1}{l}{\bfseries Site}&\multicolumn{3}{c}{\bfseries Pg\textsuperscript{\textit{a}}}&\multicolumn{1}{c}{\bfseries }&\multicolumn{3}{c}{\bfseries Rt}\tabularnewline
\cline{2-4} \cline{6-8}
\multicolumn{1}{l}{}&\multicolumn{1}{c}{Mean}&\multicolumn{1}{c}{Std. Err.}&\multicolumn{1}{c}{Anom}&\multicolumn{1}{c}{}&\multicolumn{1}{c}{Mean}&\multicolumn{1}{c}{Std. Err.}&\multicolumn{1}{c}{Anom}\tabularnewline
\hline
{\bfseries ELKVM}&&&&&&&\tabularnewline
~~Observed&8.14&0.67&0.19&&-8.19&0.69&0.21\tabularnewline
~~Filtered&3.00&0.13&0.07&&-3.06&0.14&0.07\tabularnewline
\hline
{\bfseries PDBBY}&&&&&&&\tabularnewline
~~Observed&5.95&0.69&0.22&&-5.90&0.74&0.19\tabularnewline
~~Filtered&7.01&0.40&0.05&&-7.01&0.40&0.06\tabularnewline
\hline
{\bfseries RKBMB}&&&&&&&\tabularnewline
~~Observed&3.02&0.14&0.09&&-3.62&0.15&0.08\tabularnewline
~~Filtered&3.46&0.10&0.01&&-4.07&0.11&0.01\tabularnewline
\hline
{\bfseries SAPDC}&&&&&&&\tabularnewline
~~Observed&4.89&0.23&0.13&&-6.04&0.25&0.11\tabularnewline
~~Filtered&4.94&0.09&0.00&&-6.13&0.10&0.00\tabularnewline
\hline
\end{tabular}\end{center}

>>>>

The percentage of daily integrated metabolism estimates that were anomalous (negative production, positive respiration) were significantly reduced for all sites after filtering (\cref{tab:case_res}), perhaps indicating the relative effects of water movement.  Before filtering, anomalous values ranged from 9.15 (as a percentage of the total estimates, Rookery Bay) to 21.80 (Padilla Bay) for production and 7.57 (Rookery Bay) to 20.68 (Elkhorn Slough) for respiration. Anomalous values were reduced to near zero for all case studies, particularly Rookery Bay and Sapelo Island.  Metabolism estimates using filtered \ac{DO} time series had decreased mean production (-63.1 \% change from the annual mean) and respiration (-62.6 \%) for Elkhorn Slough, increased mean production (17.8 \%) and respiration (18.8 \%) for Padilla Bay, and generally unchanged mean production and respiration for Rookery Bay and Sapelo Island (\cref{tab:case_res}).  Changes in mean estimates based on filtered \ac{DO} time series suggests that the weighted regression removed variation attributed to both biological and physical processes.  The implications of these undesirable results are described below.  Decreases in the standard deviation for all metabolism estimates were observed for all cases after filtering.  However, reduction in variation may also occur if the biological signal is reduced individually or in addition to physical variation.  The extent to which this reduction is related to the former should be minimal, provided that the two are statistically distinguishable.  Situations when the phasing of the tidal and solar cycles are correlated could be instances when the two are unable to be separated by the model.  Reductions in mean annual values (particularly at Elkhorn Slough) in addition to variance reduction suggests that true metabolism may have been removed.

@@

% case study metabolism results, including perc anom
%latex.default(tab, file = "", rowlabel = "Site", insert.bottom = foot.val,     caption = cap.val, caption.loc = "top", rgroup = unique(to_tab$Site),     n.rgroup = rep(2, 4), cgroup = c("Pg\\textsuperscript{\\textit{a}}",         "Rt"), n.cgroup = c(3, 3), rowname = rows, colheads = c(rep(c("Mean",         "SD", "Anom"), 2), c("Mean", "SD")), label = "tab:case_res")%
\begin{table}[!tbp]
\caption{Summary of metabolism estimates (gross production, respiration) for case studies using \ac{DO} time series before (observed) and after (filtered) filtering with weighted regression.  Means and standard deviation are based on daily integrated metabolism estimates.  Anomalous values are the percentage of metabolism estimates that were negative for gross production and positive for respiration.\label{tab:case_res}} 
\begin{center}
\begin{tabular}{llllclll}
\hline\hline
\multicolumn{1}{l}{\bfseries Site}&\multicolumn{3}{c}{\bfseries Pg\textsuperscript{\textit{a}}}&\multicolumn{1}{c}{\bfseries }&\multicolumn{3}{c}{\bfseries Rt}\tabularnewline
\cline{2-4} \cline{6-8}
\multicolumn{1}{l}{}&\multicolumn{1}{c}{Mean}&\multicolumn{1}{c}{SD}&\multicolumn{1}{c}{Anom}&\multicolumn{1}{c}{}&\multicolumn{1}{c}{Mean}&\multicolumn{1}{c}{SD}&\multicolumn{1}{c}{Anom}\tabularnewline
\hline
{\bfseries ELKVM}&&&&&&&\tabularnewline
~~Observed&8.14&11.44&18.64&&-8.19&11.82&20.68\tabularnewline
~~Filtered&3.00& 2.26& 6.78&&-3.06& 2.34& 7.12\tabularnewline
\hline
{\bfseries PDBBY}&&&&&&&\tabularnewline
~~Observed&5.95&11.69&21.80&&-5.90&12.60&19.03\tabularnewline
~~Filtered&7.01& 6.85& 4.84&&-7.01& 6.81& 5.54\tabularnewline
\hline
{\bfseries RKBMB}&&&&&&&\tabularnewline
~~Observed&3.02& 2.55& 9.15&&-3.62& 2.65& 7.57\tabularnewline
~~Filtered&3.46& 1.79& 0.63&&-4.07& 1.88& 0.63\tabularnewline
\hline
{\bfseries SAPDC}&&&&&&&\tabularnewline
~~Observed&4.89& 4.42&13.39&&-6.04& 4.78&10.93\tabularnewline
~~Filtered&4.94& 1.74& 0.00&&-6.13& 1.99& 0.00\tabularnewline
\hline
\end{tabular}\end{center}
